\name: A flexible and data-driven $K$-process model for nucleosynthesis

### 2023-06-18 Paper/main.tex
small hogg changes

\paragraph{6. Metallicity dependence}
We permit the process vector components for all elements other than Mg and Fe to float as a function of metallicity.
The variation is parameterized by a linear spline in log-process space, attached to a set of variable control points. We assume that a particular set of hard-coded knots are sufficient to capture the metallicity dependence. \hogg{This could be mentioned in passing in the discussion where you talk about model improvements.}

@@

However, \name{} is not without its own faults. The assumptions listed in Section~\ref{sec:model} may incorrectly skew our results and the model could benefit from improvements in implementation. While assumptions (4) and (5) on Mg and Fe production are flexible, \name{} requires that both elements have fixed process vector components. If our assumptions are incorrect and, for instance, Mg is not a pure prompt element or (in the $K=4$ case)  Fe has contributions from multiple delayed sources, our nucleosynthetic interpretation of \name{} may be incorrect. Additionally, assumption (7) states that the APOGEE data products can be used for this project, but that we inflate abundance errors with a softening parameter, $Q$ to account for their likely underestimation. If the APOGEE errors are correct, this assumption could pull the model away from a solution that favors true outlier points. In future \name{} implementation, the development of a more robust method to justifiably de-weighting outlier stars from the global fits would be beneficial. Finally, \name{} fits process vector components along a spline with 11 knots (assumption 6). As this method fits a polynomial between each knot, it can result in sharp features at the knot locations in metallicity regions with few points or large scatter. Fitting process vector components with a continuous function would be more ideal, though Appendix C.1 in G22 suggests that this change will have minimal impact on the results. 
\ejg{Hogg add a bit about expanding to a more general latent-variable model?}

@@

While such future changes will improve the model, the current form of \name{} and its data products can support ongoing research and will enable new science. Most immediately, \name{} provides high signal-to-noise abundance labels, $\Acc$ and $\AIa$, as well as de-noised stellar abundances ($m_{ij}$). $\Acc$ and $\AIa$, in particular, are powerful tracers of nucleosynthesis. \ejg{Hogg add a bit about overall $\alpha$/M indicators.} In Section~\ref{subsec:parameters}, we showed that the high-Ia and low-Ia populations are more clearly defined in $\AIa/\Acc$ vs. [Mg/H] than in [Fe/Mg] vs/ [Mg/H]. In amplitude space, the low-Ia population can be re-defined as 
\begin{equation}\label{eq:lowIa new}
\begin{cases}
\AIa/\Acc < 0.35 ,    & \mgh<-0.2 \cr
\AIa/\Acc < 0.5 + 0.7 \, \mgh,  & -0.2\geq\mgh<0.1, \cr
\AIa/\Acc < 0.57 ,    & \mgh\geq0.1 \,. \cr
\end{cases}
\end{equation}
Compared to Equation~\ref{eq:lowIa} (W19, W22), this new definition re-classifies 647 stars as high-Ia and 224 stars as low-Ia. We show the location of these stars in $\AIa/\Acc$ vs. [Mg/H] and [Mg/Fe] vs. [Fe/H] in Figure~\ref{fig:pop_divis}. Many of the re-classified stars are at $\feh>-0.1$. When dividing in [Mg/Fe], it is difficult to correctly separate the populations at high metallicity, as they are blended together. Our new definition also re-classifies many stars near [Fe/H] of -0.3 as high-Ia, suggesting that the W19 and W22 high-Ia definition has too shallow a slope. While only $\sim 2\%$ of stars are re-classified under the new definition, we suggest that Equation~\ref{eq:lowIa new} be used to chemically define the low-Ia and high-Ia populations if \name{} fits are available, especially if studying stars with $\feh > -0.1$. 
Beyond improving the definition of high-Ia and low-Ia populations, the \name{} parameters and predicted abundances could be used in any current analysis that strives to show trends with abundance labels. We predict that trends of stellar parameters with [X/H] will be clearer when comparing to $[\text{X}/\text{H}]_{\rm pred}$, $\Acc$, or $\AIa$.

>>>>

\paragraph{6. Metallicity dependence}
We permit the process vector components for all elements other than Mg and Fe to float as a function of metallicity.
The variation is parameterized by a linear spline in log-process space, attached to a set of variable control points. We assume that a particular set of hard-coded knots are sufficient to capture the metallicity dependence.

@@

However, \name{} is not without its own faults. The assumptions listed in Section~\ref{sec:model} may incorrectly skew our results and the model could benefit from improvements in implementation. While assumptions (4) and (5) on Mg and Fe production are flexible, \name{} requires that both elements have fixed process vector components. If our assumptions are incorrect and, for instance, Mg is not a pure prompt element or (in the $K=4$ case)  Fe has contributions from multiple delayed sources, our nucleosynthetic interpretation of \name{} may be incorrect. Additionally, assumption (7) states that the APOGEE data products can be used for this project, but that we inflate abundance errors with a softening parameter, $Q$ to account for their likely underestimation. If the APOGEE errors are correct, this assumption could pull the model away from a solution that favors true outlier points. In future \name{} implementation, the development of a more robust method to justifiably down-weight outlier stars from the global fits would be beneficial.
Finally, \name{} fits process vector components along a spline with 11 knots (assumption 6), and those knots have fixed locations in metallicity.
As this method fits a polynomial between each knot, it can result in sharp features at the knot locations in metallicity regions with few points or large scatter. Fitting process vector components with a differentiable function might be more reasonable, though results shown in Appendix C.1 in G22 suggests that this change might have minimal impact on the results. 
In general this model for the metallicity-dependence of the yields is very rigid; a better model could both have more flexibility and be smoother.

@@

While such future changes will improve the model, the current form of \name{} and its data products can support ongoing research and will enable new science.
Most immediately, \name{} provides high signal-to-noise abundance labels, $\Acc$ and $\AIa$, as well as de-noised stellar abundances ($m_{ij}$).
The best-fit values of $\Acc$ and $\AIa$, in particular, are powerful tracers of nucleosynthesis.
They show a bimodality at all metallicities, as do the de-noised abundances. \hogg{EJG: Is this true? I just wrote it.}
And---because the model is a maximum-likelihood model---they represent information-theory optimal combined measures of alpha  and iron abundances.
That is, these data-driven amplitudes could replace more theory-driven measures of the relative contributions of core-collapse and Ia enrichment channels.

In Section~\ref{subsec:parameters}, we showed that the high-Ia and low-Ia populations are more clearly defined in $\AIa/\Acc$ vs. [Mg/H] than in [Fe/Mg] vs/ [Mg/H]. In amplitude space, the low-Ia population can be re-defined as 
\begin{equation}\label{eq:lowIa new}
\begin{cases}
\AIa/\Acc < 0.35 ,    & \mgh<-0.2 \cr
\AIa/\Acc < 0.5 + 0.7 \, \mgh,  & -0.2\geq\mgh<0.1, \cr
\AIa/\Acc < 0.57 ,    & \mgh\geq0.1 \,. \cr
\end{cases}
\end{equation}
Compared to Equation~\ref{eq:lowIa} (W19, W22), this new definition re-classifies 647 stars as high-Ia and 224 stars as low-Ia. We show the location of these stars in $\AIa/\Acc$ vs. [Mg/H] and [Mg/Fe] vs. [Fe/H] in Figure~\ref{fig:pop_divis}. Many of the re-classified stars are at $\feh>-0.1$. When dividing in [Mg/Fe], it is difficult to correctly separate the populations at high metallicity, as they are blended together. Our new definition also re-classifies many stars near [Fe/H] of -0.3 as high-Ia, suggesting that the W19 and W22 high-Ia definition has too shallow a slope. While only $\sim 2\%$ of stars are re-classified under the new definition, we suggest that Equation~\ref{eq:lowIa new} be used to chemically define the low-Ia and high-Ia populations if \name{} fits are available, especially if studying stars with $\feh > -0.1$. 
Beyond improving the definition of high-Ia and low-Ia populations, the \name{} parameters and predicted abundances could be used in any current analysis that strives to show trends with abundance labels. We predict that trends of stellar parameters with [X/H] will be clearer when comparing to $[\text{X}/\text{H}]_{\rm pred}$, $\Acc$, or $\AIa$.
